Security Vulnerabilities in Neuroimaging Container Images

### 2020-09-10 paper.tex
Added discrepancy analysis to notebook
Fixed figure path
Rewrote discussion and conclusion

\renewcommand{\labelitemi}{$\textendash$}
\title{Security Vulnerabilities of Container Images in Neuroimaging}
\begin{filecontents*}{resul.csv}
labels,names,A,C,V,Stools
a,bids/example:latest,838,663,683,
b,bids/freesurfer:latest,353,215,192,
c,bids/ndmg:latest,934,757,782,
d,bids/broccoli:v1.0.0,203,203,117,
e,bids/fibredensityandcrosssection:latest,1086,849,868,
f,bids/spm:latest,191,246,235,
g,poldracklab/mriqc:0.15.0,432,328,332,
h,bids/cpac:v1.0.1a\_22,698,524,535,
I,bids/hyperalignment:latest,1717,1327,1433,
j,bids/mindboggle:0.0.4-1,671,955,569,
k,bids/mrtrix3\_connectome:latest,275,157,158,
l,bids/rs\_signal\_extract:0.1,867,454,473,
m,bids/aa:v0.2.0,189,244,233,
n,bids/niak:latest,1482,1366,1420,
o,bids/oppni:v0.7.0-1,913,698,723,
p,poldracklab/fmriprep:unstable,516,355,393,
q,bids/brainiak-srm:latest,897,484,503,
r,bids/nipypelines:0.3.0,641,878,451,
s,bids/hcppipelines:v3.17.0-18,526,873,620,
t,bids/magetbrain:latest,112,117,126,
u,bids/tracula:v6.0.0-4,874,664,683,
v,bids/baracus:latest,172,238,216,
w,bids/antscorticalthickness:v2.2.0-1,289,27,no data,
x,bids/dparsf:v4.3.12,507,750,757,
y,bids/afni\_proc:latest,67,16,no data,
z,bids/rshrf:1.0.1,5,3,3,
aa,bids/ndmg:v0.1.0,934,758,782,
ab,mcin/qeeg:latest,121,121,142,
ac,camarasu/creaphase:0.3,148,148,186,
ad,bigdatalabteam/hcp-prefreesurfer:exec-centos7.freesurferbuild-centos4-latest:latest,257,266,157,
ae,mcin/docker-fsl:latest,178,178,201,
af,gkiar/onevoxel:v0.3.0rc2,5,2,2,
ag,bigdatalabteam/hcp-prefreesurfer:exec-centos7-fslbuild-centos5-latest:latest,257,266,157,
ah,bt5e/ants:latest,592,600,336,
ai,poldracklab/fmriprep:1.2.3,591,441,447,
aj,gkiar/dwipreproc\_fsl-5.0.11\_minified:latest,391,258,263,
ak,gkiar/mask2boundary:v0.1.0,15,11,11,
al,mcin/ica-aroma:latest,178,178,202,
am,c3genomics/genpipes,,,,99
an,bioinformatics-group/aqua-singularity-recipe,,,,7
ao,MontrealSergiy/BEst,,,,106
ap,shub://aces/cbrain-containers-recipes:fsl\_v6.0.1,,,,206
aq,shots47s/bids-freesurfer-6.0,,,,212
ar,shots47s/bids-fmriprep-1.2.3,,,,453
\end{filecontents*}

@@

\begin{abstract}
Software containers have become key technologies for the deployment of
scientific data analyses on high-performance computing (HPC)
infrastructures, due to their positive impact on the portability, deployment, and
reproducibility of computations. Container images, however, often contain slightly outdated
software packages compared to HPC environments maintained by professional
 system administrators, and they often carry more
software packages than required by analyses, including compilers or unused
system librairies. As a result, container images are likely to contain
important amounts of security vulnerability, increasing the attack surface
for the entire infrastructure. This paper presents a vulnerability analysis
of container images for scientific data analysis. We analyze container
images with four vulnerability scanners, focusing on the use case of
neuroscience data analyses. We quantify the effect of images updates and
size reduction on the number of vulnerabilities, and conclude with
recommendations on how to build container images with a reduced amount of
vulnerabilities.

@@

We scanned all container images available at the time of this study on two containerization frameworks
used in neuroscience: BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques},
totalling
44 container images.
Out of these 44 images, 26 were obtained from BIDS apps 
and 18 from Boutiques. At the time of the study, BIDS apps had 27 images,
out of which one wasn't available on DockerHub. Boutiques had 49 images,
%listed using Boutiques \texttt{bosh search} command,
however, only 23 unique images were listed, out of which 3 couldn't be retrieved and 2
were already included in BIDS apps. All the final 26 images
from BIDS apps were Docker images, whereas Boutiques contained 18 Docker images
and 6 Singularity images.

@@

@@

\begin{figure*}
\includegraphics[width=\textwidth]{Figures/figure.pdf}
\caption{\label{fig:vulnerabilities} Number of vulnerabilities detected
by Anchore in container images. \textbf{(A)} Number of vulnerabilities
by container image and severity. \textbf{(B)} Combined
effect of image minification and package update on 5 container images. \textbf{(C, D)}
Number of vulnerabilities by number of packages, before and after
package update.}
\end{figure*}

@@

% \begin{table*}
% 	\csvreader[%
% 	 tabular={|c|c|c|c|c|c|},
% 		table head = \hline\textbf{Labels} &\textbf{{Image names}} & \textbf{Anchore} & \textbf{Clair} & \textbf{Vuls} & \textbf{Stools}\\\hline,
% 		late after line= \\,
% 			late after last line=\\\hline %
% 		]{resul.csv}{labels=\labels,names=\names,A=\A,C=\C,V=\V,Stools=\Stools}%
% 			{\labels & \names & \A & \C & \V & \Stools}
% 		   \centering
% 		\vspace*{1mm}
% 		\caption{\label{table1}Number of vulnerabilities detected by the four scanners. Vuls shows \textit{no data} for images \texttt{bids/antscorticalthickness:v2.2.0-1} and \texttt{bids/afni\_proc:latest}
% 		since their OS distributions reached end of life.} 
% 	\end{table*}
	\begin{figure}
        {\includegraphics[scale=2.5,width=\columnwidth]
        {Figures/venn.pdf}}
	\vspace*{-10mm}
	\caption{\label{fig:venn} Scanner results}
\end{figure}

@@

% files vs packages

% reproducibility vs update
However, this action affects the reproducibility
of that image. For example, if a package is used in an experiment
and later that package is updated, subsequent instances of the experiment may actually
get different results.

% minification is specific to the analysis
% tried neurodocker
% don't point at specific applications, it's a general issue

As we noticed that none of the Boutiques and BIDS images at the time of the study are free of vulnerabilities, out of which
so many are high severity vulnerabilities as well. Yet there is not a single image scanner on which we can rely upon for
scanning. Additionally, we cannot update all images because some OS releases, which are used inside images, 
already reached end-of-life and are
no longer getting updates. Furthermore, the minification of images is a time consuming process. As we have to run
the pipeline to get Reprozip trace that can take up to several hours. Also, the user should have proper knowledge
of running the pipeline because it is image-specific. In addition to that, Reprozip can be only installed
in the images having Python version 2.7.3 or greater, or 3.3 or greater. Above all, there is no guarantee that
the commands that you run to get Reprozip trace covers the whole program, maybe it runs a particular
branch of the program. So, there should be some kind of coverage measurement.
Besides this, as scanners report vulnerabilities according to the versions of the installed packages,
they are likely to miss packages that are not installed using package manager. 
Therefore, we suggest some particular rules that can be
followed to keep container images safe.
\begin{itemize}
	\item Update packages regularly inside images.
	\item Remove unnecessary packages, in particular the ones used for installation (compilers, source packages, etc.).
	\item Choose your base OS image wisely.
	\item Run image scanning tools in CI scripts. For instance Anchore.
	\item Protect your GitHub/DockerHub credentials because if automated builds is set up
in Docker Hub, it can automatically build images from external repositories, for instance GitHub.
\item Minimize dependencies inside images.
\item Avoid static linking because it obfuscates vulnerability scanning.
\item Use the package manager to facilitate vulnerability detection.
\item Use OS releases that have LTS support.
\end{itemize}
\vspace*{-2mm}
\section{Conclusions and Future Work}

In this paper, we performed various experiments to study the security state of the Docker
and Singularity images, which are used in the neuroimaging field. We found out strong correlation
between the number of packages and the number of vulnerabilities in the images, which implies
that the number of vulnerabilities increases with the number of packages. We also found out that
updating images reduces the number of vulnerabilities by a significant number. However, for some
vulnerabilities there are no fixes available yet so the images cannot 
be made totally invulnerable. Continuing with our next two experiments, we analyzed that the
image size reduction process also reduces vulnerabilities a lot. However, we found out that this is not
always true. It depends on two factors: the kind of packages and the number of packages that we are
deleting.

In future work we would like to see how these vulnerabilities can be leveraged
to design attacks from containers. Particularly, we want
to consider web servers and clusters that processes data using containers.

>>>>

\renewcommand{\labelitemi}{$\textendash$}
\title{Security Vulnerabilities in Neuroimaging Container Images}
\begin{filecontents*}{resul.csv}
labels,names,A,C,V,Stools
a,bids/example:latest,838,663,683,
b,bids/freesurfer:latest,353,215,192,
c,bids/ndmg:latest,934,757,782,
d,bids/broccoli:v1.0.0,203,203,117,
e,bids/fibredensityandcrosssection:latest,1086,849,868,
f,bids/spm:latest,191,246,235,
g,poldracklab/mriqc:0.15.0,432,328,332,
h,bids/cpac:v1.0.1a\_22,698,524,535,
I,bids/hyperalignment:latest,1717,1327,1433,
j,bids/mindboggle:0.0.4-1,671,955,569,
k,bids/mrtrix3\_connectome:latest,275,157,158,
l,bids/rs\_signal\_extract:0.1,867,454,473,
m,bids/aa:v0.2.0,189,244,233,
n,bids/niak:latest,1482,1366,1420,
o,bids/oppni:v0.7.0-1,913,698,723,
p,poldracklab/fmriprep:unstable,516,355,393,
q,bids/brainiak-srm:latest,897,484,503,
r,bids/nipypelines:0.3.0,641,878,451,
s,bids/hcppipelines:v3.17.0-18,526,873,620,
t,bids/magetbrain:latest,112,117,126,
u,bids/tracula:v6.0.0-4,874,664,683,
v,bids/baracus:latest,172,238,216,
w,bids/antscorticalthickness:v2.2.0-1,289,27,no data,
x,bids/dparsf:v4.3.12,507,750,757,
y,bids/afni\_proc:latest,67,16,no data,
z,bids/rshrf:1.0.1,5,3,3,
aa,bids/ndmg:v0.1.0,934,758,782,
ab,mcin/qeeg:latest,121,121,142,
ac,camarasu/creaphase:0.3,148,148,186,
ad,bigdatalabteam/hcp-prefreesurfer:exec-centos7.freesurferbuild-centos4-latest:latest,257,266,157,
ae,mcin/docker-fsl:latest,178,178,201,
af,gkiar/onevoxel:v0.3.0rc2,5,2,2,
ag,bigdatalabteam/hcp-prefreesurfer:exec-centos7-fslbuild-centos5-latest:latest,257,266,157,
ah,bt5e/ants:latest,592,600,336,
ai,poldracklab/fmriprep:1.2.3,591,441,447,
aj,gkiar/dwipreproc\_fsl-5.0.11\_minified:latest,391,258,263,
ak,gkiar/mask2boundary:v0.1.0,15,11,11,
al,mcin/ica-aroma:latest,178,178,202,
am,c3genomics/genpipes,,,,99
an,bioinformatics-group/aqua-singularity-recipe,,,,7
ao,MontrealSergiy/BEst,,,,106
ap,shub://aces/cbrain-containers-recipes:fsl\_v6.0.1,,,,206
aq,shots47s/bids-freesurfer-6.0,,,,212
ar,shots47s/bids-fmriprep-1.2.3,,,,453
\end{filecontents*}

@@

\begin{abstract}
Software containers have become key to the deployment of
scientific data analyses on high-performance computing (HPC)
infrastructures, due to their positive impact on the portability, deployment, and
reproducibility of computations. Container images, however, often contain slightly outdated
software packages compared to HPC environments maintained by professional
 system administrators, and they often carry more
software packages than required by analyses, including compilers or unused
system librairies. As a result, container images are likely to contain
important amounts of security vulnerability, increasing the attack surface
for the entire infrastructure. This paper presents a vulnerability analysis
of container images for scientific data analysis. We analyze container
images with four vulnerability scanners, focusing on the use case of
neuroscience data analyses. We quantify the effect of images updates and
size reduction on the number of vulnerabilities, and conclude with
recommendations on how to build container images with a reduced amount of
vulnerabilities.

@@

We scanned all container images available at the time of this study on two containerization frameworks
used in neuroscience: BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques},
totalling
44 container images.
Out of these 44 images, 26 were obtained from BIDS apps 
and 18 from Boutiques. At the time of the study, BIDS apps had 27 images,
out of which one wasn't available on DockerHub. Boutiques had 49 images,
%listed using Boutiques \texttt{bosh search} command,
however, only 23 unique images were listed, out of which 3 couldn't be retrieved and 2
were already included in BIDS apps. All the final 26 images
from BIDS apps were Docker images, whereas Boutiques contained 18 Docker images
and 6 Singularity images. \tristan{check the numbers, we have 38 images in the results}

@@

\todo{mention neuro docker}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{Figures/results.pdf}
\caption{\label{fig:vulnerabilities} Number of vulnerabilities detected
by Anchore in container images. \textbf{(A)} Number of vulnerabilities
by container image and severity. \textbf{(B)} Combined
effect of image minification and package update on 5 container images. \textbf{(C, D)}
Number of vulnerabilities by number of packages, before and after
package update.}
\end{figure*}

@@

% \begin{table*}
% 	\csvreader[%
% 	 tabular={|c|c|c|c|c|c|},
% 		table head = \hline\textbf{Labels} &\textbf{{Image names}} & \textbf{Anchore} & \textbf{Clair} & \textbf{Vuls} & \textbf{Stools}\\\hline,
% 		late after line= \\,
% 			late after last line=\\\hline %
% 		]{resul.csv}{labels=\labels,names=\names,A=\A,C=\C,V=\V,Stools=\Stools}%
% 			{\labels & \names & \A & \C & \V & \Stools}
% 		   \centering
% 		\vspace*{1mm}
% 		\caption{\label{table1}Number of vulnerabilities detected by the four scanners. Vuls shows \textit{no data} for images \texttt{bids/antscorticalthickness:v2.2.0-1} and \texttt{bids/afni\_proc:latest}
% 		since their OS distributions reached end of life.} 
% 	\end{table*}
	\begin{figure}
        \includegraphics[width=\columnwidth]{Figures/venn.pdf}
	\caption{\label{fig:venn} Comparison between image scanners.}
\end{figure}

@@

There is a widespread issue with security vulnerabilities in container
images for neuroimaging analyses, and it is likely to impact other
scientific disciplines as well. Indeed, it is common for container images
to hold hundreds of vulnerabilities, including several of critical
severity. Container images are impacted regardless of the type of analyses
that they support, and the main OS distributions Ubuntu, Debian and CentOS
are all impacted.

Software updates remove about two-thirds of the vulnerabilities found and
should certainly be considered the primary solution to this problem.
However, in neuroimaging as in other disciplines, software updates are
generally discouraged because they can affect analysis results by
introducing numerical perturbations in the
computations~\cite{gronenschild2012effects,glatard2015reproducibility}. We
believe that this position is not viable from an IT security perspective,
and that it could endanger the entire Big Data processing infrastructure,
starting with the HPC centers. Instead, we advocate a more systematic
analysis of the numerical schemes involved in data analyses, which, coupled
with software testing, would make the analyses robust to software updates.
As a first step, the packages impacting the analyses could be specifically
identified and the others updated, which would largely remove
vulnerabilities.

Ultimately, software updates should even occur at runtime rather than when
the container image is built. Indeed, it is likely that container images
used for scientific data analyses be built only occasionally, perhaps every
few weeks when a release becomes available, which may not be compatible
with the frequency of required security updates. In fact, there is no
definite reason for the application software release cycle to be
synchronized with security updates, and security updates shouldn't be
dependent on application software developers. Instead, we think it would be
relevant for analytics engines to (1) systematically apply security updates
when containers start, and (2) run software tests provided by application
developers, including numerical tests, before running analyses.

Implementing such a workflow, however, requires a long-term endeavour to
evaluate broadly the stability of data analysis pipelines, and to develop
the associated software tests. For the shorter term, we identified the
following recommendations for application developers to reduce the number
of security vulnerabilities in container images:

\begin{enumerate}[leftmargin=0pt,itemindent=*]
\item \emph{Introduce software dependencies cautiously}. Software
dependencies come with a potential security toll that is often neglected.
For instance, it can be tempting to add a complete toolbox to implement a
relatively minor operation in a data analysis pipeline, such as a data
format conversion, while the same functionality might be available in the
existing dependencies of the pipeline, albeit in a less convenient way.
\item \emph{Use lightweight base images such as Alpine Linux}. Base
images often come with packages that are useful in personnal computers or
servers, but not in containers dedicated to a specific data analysis. In
addition, lightweight distributions define packages with a fine
granularity, allowing developers to avoid installing unnecessary
dependencies. For instance, the compressed size of the Alpine Linux base
image on DockerHub is only 2.7~MB while that of the Ubuntu base image is
27~MB.
\item \emph{Use OS releases with long-term support.} Security updates are
not provided for OS distributions that reached end of life. When a given
release of a data analysis pipeline is expected to be used over a long
period of time, typically several years as it is common in neurosciences,
the life cycle of the distribution release should be considered when
choosing a base container image. OS distributions have very different life
cycle durations, as long and short life cycles serve different purposes.
For instance, among RedHat-based distributions Fedora release a new version
every 6 months and provide maintenance for about a year, while CentOS
releases every 3-5 years and provide maintenance for 10 years. Similarly,
Ubuntu LTS (long-term support) distributions provide free security updates
for 5 years, and Debian stable releases are maintained for 3 years (the
Debian LTS project intends to extend this duration to 5 years). Commercial
offers for extended maintenance periods are common. 
\item \emph{Install packages, not files}. Vulnerability scanners such as
Anchore, Clair or Vuls detect vulnerabilities from the list of packages
installed in a container image. Therefore, vulnerabilities contained in
software tools installed through direct file download rather than through
the package manager would go completely undetected. Domain-specific
distributions such as \href{http://neuro.debian.net}{NeuroDebian} or
\href{https://docs.fedoraproject.org/en-US/neurofedora/overview/}{NeuroFedora}
in neuroimaging are useful in this respect.
\item \emph{Minify container images.} The automated minification
process that we used in our study is unwieldy for a routine use, as it
requires capturing execution traces with ReproZip to reconstruct the graph
of package dependencies required for the analysis. In practice, it would be
more practical for software developers to identify and remove unnecessary
dependencies when they build containers, based on their knowledge of the
application.
\item \emph{Run image scanners during continuous integration.} Scanning
container images cann be a cumbersome process (for instance, scanning the
\todo{x} images involved in this study takes about \todo{y} hours) that
could be asynchronously executed during continuous integration (CI),
through tools such as Travis CI or Circle CI. Including security scans in
CI also allows developers to identify vulnerabilities quickly,
before new software contributions are released.
\end{enumerate}

Describing specific attacks against HPC systems that would exploit
vulnerabilities in container images is out of the scope of our study. We
believe that such attacks are likely to exist, although attacking HPC
systems through containers is challenging due to their relative isolation
from the host system. First, under the safe assumption that legitimate HPC
users can be trusted, attackers would have to be remote to the container,
either in the same network or on a remote network. Two main types of
attacks can be envisaged in these conditions: network-based attacks,
exploiting vulnerabilities in network clients installed in the container,
and data-based attacks, exploiting vulnerabilities through the processing
of malicious data injected through third-party systems.

Several types of escalation attacks could be envisaged once remote
attackers gain access to the container, in particular related to (1) using
the resources allocated to the container for malicious use, such as storing
data in the file system or using CPU cycles, resulting in denial of service
for the user running the container, and possibly for other HPC users, and
(2) attacking a host network service, for instance a scheduler or a file
system daemon. Exploits in the host kernel to break out of the container
are always possible but unlikely assuming that the host system is
maintained by professional system administrators.

\section{Conclusion}

Container images deployed on HPC clusters for scientific data analyses are
likely to contain hundreds of security vulnerabilities, many of which are
critical. In the short term, application software developers can address
this issue by: (1) minifying container images, by using lightweight OS
distributions and reducing software dependencies, and (2) applying regular
security updates, which requires using OS distributions with long-term
support. Longer term, data analysis pipelines would benefit from in-depth
stability analysis, to ensure that analytical results are not affected by
security updates.

This conclusion is not an alarming message urging HPC administrators  
to ban containers from their systems. User-controlled container images are
just one of many end-user artifacts that could serve as attack vectors,
and to our knowledge no attack has been described to exploit them. Other,
more traditional types of attacks targeting user credentials or network
connections are likely to remain more common.
